---
title: "01_Extract"
author: "Matthew Coghill"
date: "1/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document will be used to extract the spectral indices from the KXI project. First, import relevant packages.

```{r Packages, message=FALSE}

library(tidyverse)
library(sf)
library(terra)

```

Three experiments were conducted: one in Beaverdell with respect to drought, and two at the Red Lake site where hill and turning experiments were completed. As such, there are folders containing the imagery for each experiment: Beaverdell_Drought, Red_Lake_Hill, and Red_Lake_Turn. Because there are different data for each of the experiments they must be performed independently of each other. The Beaverdell site was driven on first, so let's start with that one.

First, we will set directories and import raster files for the Beaverdell drought project.

```{r Raster import}

expt <- "Beaverdell_Drought"

rast_dir <- file.path("./00_rasters", expt)
out_dir <- file.path("./01_data/Remote_Sensing")

spec_indices <- rast(list.files(rast_dir, pattern = ".tif$", full.names = TRUE))

# Clean the names of the spectral indices - Acronym of index provided after last
# underscore, so extract that and apply to each name
names(spec_indices) <- toupper(gsub("(.*_s*(.*$))", "\\2", names(spec_indices)))

dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

```

QGIS was used to create the polygons used here for raster data extraction. For the drought and hill experiments, this was straightforward: a large 35m x 2.5m rectangle was sampled from the middle 30m area by segmenting the track into 1m x 2.5m rectangles. A more involved process was used to generate the polygons for the turning experiment. The shapes are saved as geopackages which retain all column headings while being portable to many programs. One file was created for each experiment.

Import the shapes for the Beaverdell experiment into R.

```{r Shape import}

shp_dir <- file.path("./00_shapes")
lane_segs <- st_read(file.path(shp_dir, paste0("Segments_", expt, ".gpkg")), quiet = TRUE)

```

The hard work has been done in QGIS, but we still need to do a bit of data wrangling on the segments to make sure the data conforms to what we need (factors for treatments). After that, we can simply extract the raster values of each segment for each lane.

```{r Data wrangling}

lane_segs_full <- lane_segs %>% 
  dplyr::mutate(Block = factor(Block, levels = c("A", "B", "C")),
                Lane = as.integer(Lane),
                Passes = factor(Passes, levels = c(0, 1, 10, 30)),
                Vehicle = factor(Vehicle, levels = c("Control", "Standard", "KXI")),
                Water = factor(Water, levels = c("A", "W")),
                Distance = as.integer(Distance)) %>% 
  arrange(Block, Lane, Distance)

# Create a SpatVector of the above sf dataframe to use in the extraction
lane_segs_full_sv <- vect(lane_segs_full)

```

Perfect! Now we can extract raster data at each of these specified polygons and get a full suite of data. Look at means first because why not? On my laptop, this took ~2.5 minutes to run but results will vary by computer.

```{r Raster extraction}

spec_data_mean <- cbind(
  lane_segs_full,
  extract(spec_indices, lane_segs_full_sv, fun = mean, na.rm = TRUE, ID = FALSE, exact = TRUE) %>% 
    data.frame())

st_drop_geometry(spec_data_mean) %>% 
  write.csv(file.path(out_dir, paste0(expt, "_mean_index_values.csv")), row.names = FALSE)

```

That's it for this experiment. Now, repeat these steps for the other 2 experiments that occurred in Red Lake.

The methods are pretty much replicated for the hill experiment (the middle 30m of the 35m long lane is sampled for each lane in 1m increments).

```{r Red Lake Hills experiment}

expt <- "Red_Lake_Hills"
rast_dir <- file.path("./00_rasters", expt)
shp_dir <- file.path("./00_shapes")
spec_indices <- rast(list.files(rast_dir, pattern = ".tif$", full.names = TRUE))

# Clean the names of the spectral indices - Acronym of index provided after last
# underscore, so extract that and apply to each name
names(spec_indices) <- toupper(gsub("(.*_s*(.*$))", "\\2", names(spec_indices)))

lane_segs <- st_read(file.path(shp_dir, paste0("Segments_", expt, ".gpkg")), quiet = TRUE)
lane_segs_full <- lane_segs %>% 
  dplyr::mutate(Block = factor(Block, levels = c("A", "B", "C")),
                Lane = as.integer(Lane),
                Vehicle = factor(Vehicle, levels = c("Control", "Standard", "KXI")),
                Distance = as.integer(Distance)) %>% 
  arrange(Block, Lane, Distance)

# Create a SpatVector of the above sf dataframe to use in the extraction
lane_segs_full_sv <- vect(lane_segs_full)

spec_data_mean <- cbind(
  lane_segs_full,
  extract(spec_indices, lane_segs_full_sv, fun = mean, na.rm = TRUE, ID = FALSE, exact = TRUE) %>% 
    data.frame())

st_drop_geometry(spec_data_mean) %>% 
  write.csv(file.path(out_dir, paste0(expt, "_mean_index_values.csv")), row.names = FALSE)

```

The turning experiment can be investigated a bit more thoroughly if needed. Polygons within a turn have a short side and a long side, allowing a given polygon to be designated as "right" or "left" turning based on these distances. Distances are provided with respect to the side of the truck (i.e.: the column "left_length" refers to the length of the polygon on the left side of the truck). If a distance on the left side is less than the distance on the right side of the truck, that would indicate a left turn for that polygon. Some amount of tolerance should be applied to allow for minute differences in left vs. right lengths (perhaps 0.1m?). I will include extra lines in the "mutate()" function to classify such turns.

```{r Red Lake Turns experiment}

expt <- "Red_Lake_Turns"
rast_dir <- file.path("./00_rasters", expt)
shp_dir <- file.path("./00_shapes")
spec_indices <- rast(list.files(rast_dir, pattern = ".tif$", full.names = TRUE))

# Clean the names of the spectral indices - Acronym of index provided after last
# underscore, so extract that and apply to each name
names(spec_indices) <- toupper(gsub("(.*_s*(.*$))", "\\2", names(spec_indices)))

lane_segs <- st_read(file.path(shp_dir, paste0("Segments_", expt, ".gpkg")), quiet = TRUE)
lane_segs_full <- lane_segs %>% 
  dplyr::mutate(Block = factor(Block, levels = c("A", "B", "C")),
                Lane = as.integer(Lane),
                Passes = factor(Passes, levels = c(0, 2, 4)),
                Vehicle = factor(Vehicle, levels = c("Control", "Standard", "KXI")),
                Distance = as.integer(Distance),
                Turn = factor(case_when(
                  left_length - right_length > 0.1 ~ "Right",
                  right_length - left_length > 0.1 ~ "Left",
                  TRUE ~ "Straight"
                ), levels = c("Straight", "Left", "Right"))) %>% 
  arrange(Block, Lane, Distance)

# Create a SpatVector of the above sf dataframe to use in the extraction
lane_segs_full_sv <- vect(lane_segs_full)

spec_data_mean <- cbind(
  lane_segs_full,
  extract(spec_indices, lane_segs_full_sv, fun = mean, na.rm = TRUE, ID = FALSE, exact = TRUE) %>% 
    data.frame()) %>% 
  dplyr::select(-c(right_length, left_length))

st_drop_geometry(spec_data_mean) %>% 
  write.csv(file.path(out_dir, paste0(expt, "_mean_index_values.csv")), row.names = FALSE)

```
